---
title: "MPSA 2026 - estimation results for paper"
author: "Sofía Granados and Linette Page"
date: "2026-04-19"
output: pdf_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

# 1. Set up

```{r load data and packages, warning=FALSE, include=FALSE, message=FALSE}
library(tidyverse)
library(estimatr)
library(Matching)
library(ebal)
library(kableExtra)
library(modelsummary)
library(stats)
library(MatchIt)
library(WeightIt)
library(gridExtra)
#install.packages("devtools")
#devtools::install_github("chadhazlett/kbal") #installing chad's kbal
library(kbal)
library(sensemakr)
library(RColorBrewer)
library("cobalt")

# Working directory for all to work - Comment as needed!!!!
#setwd("/Users/sofiagranados/Library/CloudStorage/Dropbox/Page and Granados") # For SG
# full_dataset <- read_csv("data/out/full_dataset.csv")
# estimation_data <- read_csv("data/out/estimation_data.csv")

full_dataset <- read_csv("../data/out/full_dataset.csv") #linette
estimation_data <- read_csv("../data/out/estimation_data.csv") #linette
```

# 2. Data cleaning

Treatment is disappearances and homicides by the state; calculated as the yearly rate by 10,000

## 2.1 choosing treatment and covariates

```{r treat_definition}
#select treatment
estimation_data$treat <- estimation_data$etv_dh
#other violence data options: $etv_all_jep, $etv_cnmh, $etv_all_cnmh, $etv_false_pos

#versions of treatment
estimation_data <- estimation_data |>
  mutate(treat_type = ifelse(treat > 0, 1, 0))

estimation_nona <- estimation_data |>
  drop_na() #1115 obs to 1006 - which ones?; are the missing values balanced?

#covariates
covariates <- c("uribe_2002_proportion", "log_pop", "gdp_pcap_cons2005", "gini", "etv_state_pre", "etv_guerr_pre", "hist_land_conflict", "justice_percent", "rank_w_any")

balance_formula <- as.formula(paste("treat_type ~ ", paste(covariates, collapse=" + ")))

```


### 2.2 Covariates
A density plot of the yearly rate of cases per 10,000 per municipality; 
followed by the exact number and proportion of control municipalities

```{r treatment cutoff}
estimation_nona |>
  ggplot(aes(x = treat)) + #spread of etv proportion
    geom_density()

table(estimation_nona$treat)[1] #number of zero state violence
table(estimation_nona$treat)[[1]]/nrow(estimation_nona) #percent of municipalities that have not experienced any state violence
```

#### 2.2.1 Covariate balance

```{r balance plot before, fig.align='center'}

estimation_nona |>
  dplyr::select(all_of(covariates), treat_type) |> #choose covariates
  reshape2::melt(all_of(covariates), id.vars = "treat_type", na.rm = T) |> #put them all in one column
  ggplot(aes(x = value)) +
  geom_density(aes(col = as_factor(treat_type)), alpha = 0.4) +
  facet_wrap(~ variable, scales = "free") +
  labs(title = paste0("PDF of Covariates by Treatment Status"),
       y = "Density", col = "Treatment") +
  labs(col = "Treatment") +
  theme_minimal() +
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank()) +
  scale_color_manual(values = c("darkorange2", "orchid3"), labels = c("Control", "Treatment")) 
```

```{r balance table before}

match_info <- MatchBalance(balance_formula, data = estimation_nona, nboots = 1000, print.level = 0) #checking balance before matching

balance_test_before <- baltest.collect(match_info, var.names = covariates, after = FALSE) #gets numbers from match_info
balance_test_before <- balance_test_before[, c("mean.Tr", "mean.Co", "T pval", "KS pval")] #rename columns

balance_test_before |> #output table
  kable(booktabs = T, digits = 3, align = rep('c', 4),
        col.names = c("Treatment Mean", "Control Mean", "T p-value", "KS p-value")) |>
  kable_styling(full_width = F, position = "center")
```



# 3 Versions of matching and weighting

## 3.1 Naive regression

```{r naive}
lm_naive <- lm_robust(trust_plebi_yes ~ treat_type, data = estimation_nona)
```

## 3.2 Naive + Control regression

```{r controls}

lm_controls <- lm_lin(trust_plebi_yes ~ treat_type, covariates = ~ uribe_2002_proportion + log_pop + gdp_pcap_cons2005 +  gini + etv_state_pre + etv_guerr_pre + hist_land_conflict + justice_percent + rank_w_any, data = estimation_nona) 
```

## 3.3 Matching - covariates

Throws away non-matched control units

- No replacement and one-to-one matching
- with Match(): closest $X_i$ through normalized Euclidean distance
- BiasAdjust = TRUE: bias correction to adjust for multiple continuous variables
- other things to consider: how many matches, which algorithm (eg Coarse, pre-processing), whether to drop treatment units to remove bias, calipers to remove units with no common support


First, creating a matched dataset:

```{r matching}
exact_matches <- ifelse(colnames(estimation_nona[,covariates]) %in% c("hist_land_conflict"), TRUE, FALSE) #distinguish between continuous and categorical (which you can exact match on)

match_model_1 <- Match(Y = estimation_nona$trust_plebi_yes,
                       Tr = estimation_nona$treat_type,
                       X = estimation_nona[, covariates],
                       M = 1, # one-to-one matching
                       exact = exact_matches,
                       caliper = 2, #0.25 standard deviations dropped
                       Weight = 2, #inverse of variance
                       replace = TRUE,
                       estimand = "ATT", 
                       BiasAdjust = TRUE) # bias adjustment

#get dataset of just matched observations
match_data_1 <- estimation_nona[c(match_model_1$index.treated, 
                                  match_model_1$index.control),]
table(match_data_1$treat_type)
```

```{r run matching again, warning=FALSE, include=FALSE, message=FALSE}
# #run again - doesn't make sense to do
# match_model_2 <- Match(Y = match_data_1$trust_plebi_yes,
#                        Tr = match_data_1$treat_type,
#                        X = match_data_1[, covariates],
#                        M = 1, # one-to-one matching
#                        estimand = "ATT", 
#                        BiasAdjust = TRUE) # bias adjustment
# 
# match_data_2 <- estimation_nona[c(match_model_2$index.treated, 
#                                   match_model_2$index.control),]
# 
# match_data_2 <- match_data_2 |>
#   drop_na()

```
Should we run it multiple times? The results don't change much

Then, checking the balance after the matching

```{r balance table after}
match_info <- MatchBalance(match.out = match_model_1, #same as above - gets balance, but this time compares before and after
                                    formul = balance_formula,
                                    data = estimation_nona,
                                    print.level = 0, nboots = 1000)

balance_test_after <- baltest.collect(match_info, var.names = covariates, after = TRUE) #new balance test
balance_test_after <- balance_test_after[, c("mean.Tr", "mean.Co", "T pval", "KS pval")]

balance_test_after %>%
  kable(booktabs = T, digits = 3, align = rep('c', 4),
        col.names = c("Treatment Mean", "Control Mean", "T p-value", "KS p-value")) %>% 
  kable_styling(full_width = F, position = "center")
```

```{r second run balance table, warning=FALSE, include=FALSE, message=FALSE}
# #second round - the balance table looks the same as the first round
# match_info_1 <- MatchBalance(match.out = match_model_2,
#                                     formul = balance_formula,
#                                     data = match_data_1,
#                                     print.level = 0, nboots = 1000)
# 
# match_info_1 <- MatchBalance(balance_formula,
#                                     data = match_data_2,
#                                     print.level = 0, nboots = 1000)
# 
# balance_test_after_1 <- baltest.collect(match_info_1, var.names = covariates, after = FALSE)
# balance_test_after_1 <- balance_test_after[, c("mean.Tr", "mean.Co", "T pval", "KS pval")]
# balance_test_after_1 %>%
#   kable(booktabs = T, digits = 3, align = rep('c', 4),
#         col.names = c("Treatment Mean", "Control Mean", "T p-value", "KS p-value")) %>% 
#   kable_styling(full_width = F, position = "center")
```

```{r balance plot after}

match_data_1 |>
  dplyr::select(all_of(covariates), treat_type) |> #choose covariates
  reshape2::melt(all_of(covariates), id.vars = "treat_type", na.rm = T) |> #put them all in one column
  ggplot(aes(x = value)) +
  geom_density(aes(col = as_factor(treat_type)), alpha = 0.4) +
  facet_wrap(~ variable, scales = "free") +
  labs(title = paste0("PDF of Covariates by Treatment Status, after matching on covariates"),
       y = "Density", col = "Treatment") +
  labs(col = "Treatment") +
  theme_minimal() +
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank()) +
  scale_color_manual(values = c("darkorange2", "orchid3"), labels = c("Control", "Treatment")) 
```

Finally, calculating ATT
```{r calculate att}

lm_match_covariates <- lm_lin(trust_plebi_yes ~ treat_type, covariates = ~ uribe_2002_proportion + log_pop + gdp_pcap_cons2005 +  gini + etv_state_pre + etv_guerr_pre + hist_land_conflict + justice_percent + rank_w_any, data = match_data_1)

```


```{r remove 2, warning=FALSE, include=FALSE, message=FALSE}

rm(balance_test_before, balance_test_after, match_info, match_data_1, match_model_1,  exact_matches, lm_match, lm_match_controls)
```

## 3.4 Matching - Propensity score matching

```{r compare propensity scores}
# calculate propensity scores
estimation_nona$ps <- stats::glm(balance_formula, 
                                 data = estimation_nona, 
                                 family = binomial(link = logit)
                                 )$fitted.values

ps = estimation_nona$ps
D = estimation_nona$treat_type

#plot propensity scores of treated vs control
ggplot(tibble(ps = estimation_nona$ps,
              treatment = as_factor(ifelse(estimation_nona$treat_type == 1, "Treated", "Control"))),
       aes(x = ps, col = treatment)) +
  geom_density() +
  labs(title = "Distribution of propensity scores before", y = "Density", x = "Propensity Score") +
  xlim(c(0, 1)) +
  ylim(c(0, 5)) +
  theme_minimal() +
  scale_color_manual(values = c("darkorange2", "orchid3"), labels = c("Treatment", "Control"))

```

Compare units that have a comparable probability of getting treated to match:

```{r match propensity scores}

#match based on propensity scores
match_ps <- Match(Y = estimation_nona$trust_plebi_yes, 
                  Tr = estimation_nona$treat_type, 
                  X = estimation_nona$ps, 
                  M = 1, 
                  estimand = "ATT")

#get dataset of only matched observations
match_data_ps <- estimation_nona[c(match_ps$index.treated, 
                                  match_ps$index.control),]
table(match_data_ps$treat_type)


```

Checking balance after matching:

```{r balance after 2}
#get balance of dataset after matching
match_balance_ps <- MatchBalance(match.out = match_ps, 
                                 formul = balance_formula, 
                                 data = estimation_nona, 
                                 print.level = 0)

balance_table_ps <- baltest.collect(match_balance_ps, var.names = covariates, after = TRUE) #collecting numbers from matchbalance
balance_table_ps <- balance_table_ps[, c("mean.Tr", "mean.Co", "T pval", "KS pval")]

balance_table_ps |> #output table
  kable(booktabs = T, digits = 3, align = rep('c', 4),
        col.names = c("Treatment Mean", "Control Mean", "T p-value", "KS p-value")) |> 
  kable_styling(full_width = F, position = "center")

#plot of propensity scores after matching
ggplot(tibble(ps = match_data_ps$ps, 
              treatment = as_factor(ifelse(match_data_ps$treat_type == 1, "Treated", "Control"))),
       aes(x = ps, col = treatment)) +
  geom_density() +
  labs(title = "Distribution of propensity scores after matching", y = "Density", x = "Propensity Score") +
  xlim(c(0, 1)) +
  theme_minimal() +
  scale_color_manual(values = c("darkorange2", "orchid3"), labels = c("Control", "Treatment")) 
```

Calculating the ATT:
```{r calculate att 2}

lm_match_ps <- lm_lin(trust_plebi_yes ~ treat_type, covariates = ~ uribe_2002_proportion + log_pop + gdp_pcap_cons2005 +  gini + etv_state_pre + etv_guerr_pre + hist_land_conflict + justice_percent + rank_w_any, data = match_data_ps)
```

The Match() propensity scores afterwards are very close, but the data has been significantly manipulated: there are many duplicates of both control and treatment

## 3.5 Weighting - propensity scores

Inverse Probability-of-Treatment Weighting (IPW):

```{r weight propensity scores}
# stabilized IPW weights, created with propensity scores
ps = estimation_nona$ps
D = estimation_nona$treat_type
PrD = mean(D)
IPW = (D*PrD+(1-D)*(1-PrD))/(D*ps+(1-D)*(1-ps))

# check the propensity scores before weighting
plot(density(ps[D==1]), lwd = 2, ylim = c(0, 4.5), 
     main="Distribution of propensity scores: before weighting")
lines(density(ps[D==0]),lwd = 2, lty = 2)
legend("topleft", legend = c("treated","controls"), lty = c(1, 2), lwd = 2)

# check the propensity scores after weighting
plot(density(ps[D==1], weight = IPW[D==1]/sum(IPW[D==1])), lwd = 2, ylim = c(0, 4.5), 
     main="Distribution of propensity scores: Weighted")
lines(density(ps[D==0], weight = IPW[D==0]/sum(IPW[D==0])),lwd = 2, lty = 2)
legend("topleft", legend = c("treated","controls"), lty = c(1, 2), lwd = 2)

IPW_df <- as.data.frame(IPW)

IPW_df |>
  ggplot(aes(x = IPW)) +
    geom_density() +
    labs(
    title = "Density plot of IPW",
    x = "weight"
  ) +
  theme_minimal() #check IPW for extremes 

```

Check how extreme the weights get / how few units are doing most of the work - it means large bias

Take regressions:
```{r regression with weights}

lm_weight_ps <- lm_lin(trust_plebi_yes ~ treat_type, covariates = ~ uribe_2002_proportion + log_pop + gdp_pcap_cons2005 +  gini + etv_state_pre + etv_guerr_pre + hist_land_conflict + justice_percent + rank_w_any, weight = IPW, data = estimation_nona)
```

## 3.6 Weighting - Kernel balancing weights

```{r kernel}
# https://cran.r-project.org/web/packages/KRLS/KRLS.pdf

weight_model_kbal = kbal(allx = estimation_nona[, covariates],
                         treatment = estimation_nona$treat_type,
                         printprogress = FALSE)

#adding weights to main dataset
estimation_nona$kernel_weights <- weight_model_kbal$w

#lin regression
lm_weight_ker <- lm_lin(trust_plebi_yes ~ treat_type, covariates = ~ uribe_2002_proportion + log_pop + gdp_pcap_cons2005 + gini + etv_state_pre + etv_guerr_pre + hist_land_conflict + justice_percent + rank_w_any, weights = weight_model_kbal$w, data = estimation_nona)

```

Propensity scores before and after kernel weighting
```{r kernel ps, fig.align='center'}

# check the propensity scores before weighting
plot(density(ps[D==1]), lwd = 2, ylim = c(0, 4.5), 
     main="Distribution of propensity scores: before weighting")
lines(density(ps[D==0]),lwd = 2, lty = 2)
legend("topleft", legend = c("treated","controls"), lty = c(1, 2), lwd = 2)


# check the propensity scores after weighting
plot(density(ps[D==1], weight = weight_model_kbal$w[D==1]/sum(weight_model_kbal$w[D==1])), lwd = 2, ylim = c(0, 4.5), 
     main="Distribution of propensity scores: Kernel weighted")
lines(density(ps[D==0], weight = weight_model_kbal$w[D==0]/sum(weight_model_kbal$w[D==0])),lwd = 2, lty = 2)
legend("topleft", legend = c("treated","controls"), lty = c(1, 2), lwd = 2)

```

Balance tables, before and after kernel weighting
```{r kernel balance tables}

# balance table before weighting

bal.tab(balance_formula, data = estimation_nona,
        disp = c("means"), un = FALSE, 
        stats = c("mean.diffs"))

# balance table after weighting

bal.tab(balance_formula, data = estimation_nona,
        weights = weight_model_kbal$w,
        disp = c("means"), un = FALSE, 
        stats = c("mean.diffs"))
```

```{r}
#shows level of bias bound at each K; kbal() chooses where it is the lowest
plot(x = weight_model_kbal$dist_record[1,],
     y = sd(estimation_nona$trust_plebi_yes) * sqrt(55) * weight_model_kbal$dist_record[2,],
     ylab="biasbound", 
     xlab="Num. dims of K balanced",
     pch=16,
     main = "Dimensons of K and biasbound")

ker_df <- as.data.frame(weight_model_kbal[["w"]])

ker_df |>
  ggplot(aes(x = weight_model_kbal[["w"]])) +
    geom_density() +
    labs(
    title = "Density plot of kernel weights",
    x = "weight"
  ) +
  theme_minimal()  +
  scale_color_manual(values = c("darkorange2", "orchid3"), labels = c("Control", "Treatment"))

```

Where treatment and control are the most balanced (ebalance convergence) and bias bound is as low as possible


```{r remove 3, warning=FALSE, include=FALSE, message=FALSE}
rm(balance_table_ps, match_ps, match_balance_ps, match_data_ps,  weight_model_kbal, weight_model_ps, D, IPW, PrD, ps, lm_match_ps_controls, IPW_df, ker_df)
```


# 5. Results

```{r all models table - with control and naive}

#all the models
models <- list(
  "(1)" = lm_naive,
  "(2)" = lm_controls,
  "(3)" = lm_match_covariates,
  "(4)" = lm_match_ps,
  "(5)" = lm_weight_ps,
  "(6)" = lm_weight_ker
  )


#put together
modelsummary(models, 
             #vcov = "HC2",
             #statistic = "p.value",
             stars = TRUE,
             coef_rename = c("treat_type" = "treatment", "uribe_2002_proportion_c" = "uribe vote", "log_pop_c" = "population", "gdp_pcap_cons2005_c" = "gdp per capita", "gini_c" = "gini", "etv_state_pre_c" = "state violence", "etv_guerr_pre_c" = "guerilla violence", "hist_land_conflict_c" = "historical land conflict", "justice_percent_c" = "judicial capacity", "rank_w_any_c" = "military rank"),
             coef_omit = ":",
             gof_omit = "BIC|AIC|Log.Lik|F",
             title = "The effect of state violence cases on support for peace",
             escape = FALSE, #to fix bug in header; need to include "+ p < 0.1, * p < 0.05, ** p < 0.01, *** p < 0.001" somewhere
             output = "kableExtra") |> 
  kable_styling(full_width = TRUE, position = "center", font_size = 8, latex_options = "HOLD_position") |>
  add_header_above(c(" " = 1, "Naive" = 1, "Controls" = 1, "Covariates" = 1, "Propensity scores" = 2, "Kernel" = 1)) |>
  add_header_above(c(" " = 3, "Matching" = 2, "Weighting" = 2)) #|>
  #save_kable("code/outputs/falsos_table.png")

```

```{r all models plot}

models_plot <- list(
  "naive" = lm_naive,
  "controls" = lm_controls,
  "covariate matching" = lm_match_covariates,
  "propensity score matching" = lm_match_ps,
  "propensity score weighting" = lm_weight_ps,
  "kernel weighting" = lm_weight_ker
  )

#plot of just treat_type variable
modelplot(models_plot,
          facet=TRUE,
          coef_omit = c(-2)) +
  aes(color = model, shape = model) +
  scale_color_manual(values = c("grey", "grey", "darkorange2", "orange1", "slateblue2", "slateblue4")) +
  scale_shape_manual(values = c(1, 1, 1, 1, 1, 1)) +
  theme(legend.position="none") +
  geom_vline(xintercept = 0) +
  labs(
    title = "The effect of state violence cases on support for peace",
    x = "coefficient estimates of proportion of yes votes per municipality",
    y = "dichotomized state violence cases",
  )

```


# 6. Sensitivity analysis

Which covariate as a benchmark?

```{r sensitivity}

estimation_model <- lm(trust_plebi_yes ~ treat_type + uribe_2002_proportion + log_pop + gdp_pcap_cons2005 + gini + etv_state_pre + etv_guerr_pre + hist_land_conflict + justice_percent + rank_w_any, data = estimation_nona)


estimation_sensitivity <- sensemakr(model = estimation_model, 
                                treatment = "treat_type",
                                benchmark_covariates = "gini",
                                  #list(covariates = c("uribe_2002_proportion", "log_pop", "gdp_pcap_cons2005", "gini", "etv_state_pre", "etv_guerr_pre", "hist_land_conflict", "justice_percent", "rank_w_any")),
                                kd = c(1:3), #1-3 times stronger than treatment
                                ky = c(1:3), #1-3 times stronger than outcome
                                q = 1, #reduction of 100% of current estimate - true effect of zero is problematic
                                alpha = 0.05, 
                                reduce = T)

#ovb_minimal_reporting(estimation_sensitivity, format = "latex")

#summary(estimation_sensitivity)

plot(estimation_sensitivity)
```

Sensitivity with regular lm() only; ipw/matching weights mess it up -- how to add other versions?

```{r remove 4, warning=FALSE, include=FALSE, message=FALSE}
rm(estimation_model, estimation_sensitivity)
```
