---
title: "200C - Project Estimation"
author: "Sofía Granados and Linette Page"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---
Last edited: Jun 8 2025 2:30am LP

# 1. Set up

```{r load data and packages, warning=FALSE, include=FALSE, message=FALSE}
library(tidyverse)
library(estimatr)
library(Matching)
library(ebal)
library(kableExtra)
library(modelsummary)
library(stats)
library(MatchIt)
library(WeightIt)
#install.packages("devtools")
#devtools::install_github("chadhazlett/kbal") #installing chad's kbal
library(kbal)
library(sensemakr)
library(RColorBrewer)
library("cobalt")

# Working directory for all to work - Comment as needed!!!!
#setwd("/Users/sofiagranados/Library/CloudStorage/Dropbox/Page and Granados") # For SG
#setwd("~/Library/CloudStorage/Dropbox/Page and Granados") # For LP

full_dataset <- read_csv("data/out/full_dataset.csv")
estimation_data <- read_csv("data/out/estimation_data.csv")
```

# 2. State-based violence (Dissappearences and homicides by the state)

## 2.1 Data cleaning 

```{r treat_definition}
#select treatment
estimation_data$treat <- estimation_data$etv_dh
#estimation_data$treat <- estimation_data$etv_all_jep
#estimation_data$treat <- estimation_data$etv_cnmh
#estimation_data$treat <- estimation_data$etv_all_cnmh
#estimation_data$treat <- estimation_data$etv_false_pos

#dichotomized treatment
estimation_data <- estimation_data |>
  mutate(treat_dicho = ifelse(treat > 0, 1, 0))

#missings
#TODO: NA's found in data input.  IT IS HIGHLY RECOMMENDED THAT YOU TEST IF THE MISSING VALUES ARE BALANCED INSTEAD OF JUST DELETING THEM.

estimation_data |>
  summarise_all(~ sum(is.na(.))) #how many missing for each

estimation_nona <- estimation_data |>
  drop_na() #1115 obs to 1006 - which ones?

#covariates
covariates <- c("uribe_2002_proportion", "log_pop", "gdp_pcap_cons2005", "gini", "etv_state_pre", "etv_guerr_pre", "hist_land_conflict")
#dropped: "uribe_2006_proportion", "avg_schooling", "literacy", "prop_rural", "mpi", "dem_security",
#all: covariates <- c("uribe_2002_proportion", "uribe_2006_proportion", "avg_schooling", "literacy", "log_pop", "prop_rural", "gdp_pcap_cons2005", "mpi", "gini", "etv_state_pre", "etv_nonstate_pre", "dem_security", "hist_land_conflict")


balance_formula <- as.formula(paste("treat_dicho ~ ", paste(covariates, collapse=" + ")))

# calculate propensity scores
estimation_nona$ps <- stats::glm(balance_formula, 
                                 data = estimation_nona, 
                                 family = binomial(link = logit)
                                 )$fitted.values
```

## 2.2 Naive regression

Estimated treatment effect:
```{r naive}
#naive <- lm_robust(trust_plebi_yes ~ treat, data = estimation_nona) #continuous
lm_naive <- lm_robust(trust_plebi_yes ~ treat_dicho, data = estimation_nona) #dichotomized

```


## 2.3 Naive + Control regression

```{r controls}

#lm_robust(trust_plebi_yes ~ treat + uribe_2002_proportion + uribe_2006_proportion + avg_schooling + literacy + log_pop + prop_rural + gdp_pcap_cons2005 + mpi + gini + etv_state_pre + etv_nonstate_pre + dem_security + hist_land_conflict, data = estimation_nona) #continuous

lm_controls <- lm_robust(trust_plebi_yes ~ treat_dicho + uribe_2002_proportion + log_pop + gdp_pcap_cons2005 +  gini + etv_state_pre + etv_guerr_pre + hist_land_conflict, data = estimation_nona) #dicotomized
#all: trust_plebi_yes ~ treat_dicho + uribe_2002_proportion + uribe_2006_proportion + avg_schooling + literacy + log_pop + prop_rural + gdp_pcap_cons2005 + mpi + gini + etv_state_pre + etv_nonstate_pre + dem_security + hist_land_conflict, data = estimation_nona
```

## 2.4. Matching

### 3.1 Covariates

```{r treatment cutoff}
estimation_nona |>
  ggplot(aes(x = treat)) + #spread of etv proportion
    geom_density()

table(estimation_nona$treat)[1] #number of zero state violence
table(estimation_nona$treat)[[1]]/nrow(estimation_nona) #percent of municipalities that have not experienced any state violence
```


#### 3.1.1 Covariate balance
```{r balance plot}

estimation_nona |>
  dplyr::select(all_of(covariates), treat_dicho) |> #choose covariates
  reshape2::melt(all_of(covariates), id.vars = "treat_dicho", na.rm = T) |> #put them all in one column
  ggplot(aes(x = value)) +
  geom_density(aes(col = as_factor(treat_dicho)), alpha = 0.4) +
  facet_wrap(~ variable, scales = "free") +
  labs(title = paste0("PDF of Covariates by Treatment Status"),
       y = "Density", col = "Treatment") +
  labs(col = "Treatment") +
  theme_minimal() +
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank()) +
  scale_color_manual(values = c("darkorange2", "orchid3"), labels = c("Control", "Treatment")) 
```


```{r balance table}

match_info <- MatchBalance(balance_formula, data = estimation_nona, nboots = 1000, print.level = 0) #checking balance before matching

balance_test_before <- baltest.collect(match_info, var.names = covariates, after = FALSE) #gets numbers from match_info
balance_test_before <- balance_test_before[, c("mean.Tr", "mean.Co", "T pval", "KS pval")] #rename columns

balance_test_before |> #output table
  kable(booktabs = T, digits = 3, align = rep('c', 4),
        col.names = c("Treatment Mean", "Control Mean", "T p-value", "KS p-value")) |>
  kable_styling(full_width = F, position = "center")
```

These are almost all significantly different between treatment and control.

#### 3.1.2 Matching covariates

Matching throws away non-matched control units

-Replacement = False
- how many matches: one to one


TODO: Why ATT? We are grabbing the control unit with the closest $X_i$
What does closest mean? For Match(): normalized Euclidean distance

curse of dimensionality: the more covariates, the further
  Todo: should we think through each covariate and whether we really need it?
    Also: issue of overspecifying the model
    - run model treatment ~ covariates
  
Issue of discrepancy with multiple continuous variables
  Have to adjust for it (bias correction) through BiasAdjust = TRUE
  Subtracts off estimated discrepancy
  
TODO: Different combinations: how many matches; which algorithm (Coarse - too many covariates means too far distance; pre-processing - match then regression on only close); whether to drop treatment units (removes bias)

caliper: removes units with no common support


```{r matching}
exact_matches <- ifelse(colnames(estimation_nona[,covariates]) %in% c("dem_security", "hist_land_conflict"), TRUE, FALSE) #distinguish between continuous and categorical (which you can exact match on)

match_model_1 <- Match(Y = estimation_nona$trust_plebi_yes,
                       Tr = estimation_nona$treat_dicho,
                       X = estimation_nona[, covariates],
                       M = 1, # one-to-one matching
                       exact = exact_matches,
                       caliper = 2, #0.25 standard deviations dropped
                       Weight = 2, #inverse of variance
                       replace = TRUE,
                       estimand = "ATT", 
                       BiasAdjust = TRUE) # bias adjustment

#get dataset of just matched observations
match_data_1 <- estimation_nona[c(match_model_1$index.treated, 
                                  match_model_1$index.control),]
table(match_data_1$treat_dicho) #596 treated, 596 control

# #run again - doesn't make sense to do
# match_model_2 <- Match(Y = match_data_1$trust_plebi_yes,
#                        Tr = match_data_1$treat_dicho,
#                        X = match_data_1[, covariates],
#                        M = 1, # one-to-one matching
#                        estimand = "ATT", 
#                        BiasAdjust = TRUE) # bias adjustment
# 
# match_data_2 <- estimation_nona[c(match_model_2$index.treated, 
#                                   match_model_2$index.control),]
# 
# match_data_2 <- match_data_2 |>
#   drop_na()

```

```{r calculate att}
match_model_1$est
match_model_1$se
summary(match_model_1)
# Estimate...  -0.011164 
# AI SE......  0.0077703 
# T-stat.....  -1.4367 
# p.val......  0.1508

#calculate ATT manually
mean(match_data_1$trust_plebi_yes[match_data_1$treat_dicho == 1]) - mean(match_data_1$trust_plebi_yes[match_data_1$treat_dicho == 0]) #0.005162898

#naive matching
#lm_match <- lm_robust(trust_plebi_yes ~ treat_dicho, data = match_data_1) #0.005162898

#controls matching
#lm_match_controls <- lm_robust(trust_plebi_yes ~ treat_dicho + uribe_2002_proportion + log_pop + gdp_pcap_cons2005 +  gini + etv_state_pre + etv_nonstate_pre + hist_land_conflict, data = match_data_1) #-0.0126571514

```


#### 3.1.3 Checking the balance after the matching

```{r balance after}
match_info <- MatchBalance(match.out = match_model_1, #same as above - gets balance, but this time comparesbefore and after
                                    formul = balance_formula,
                                    data = estimation_nona,
                                    print.level = 0, nboots = 1000)

balance_test_after <- baltest.collect(match_info, var.names = covariates, after = TRUE) #new balance test
balance_test_after <- balance_test_after[, c("mean.Tr", "mean.Co", "T pval", "KS pval")]

balance_test_after %>%
  kable(booktabs = T, digits = 3, align = rep('c', 4),
        col.names = c("Treatment Mean", "Control Mean", "T p-value", "KS p-value")) %>% 
  kable_styling(full_width = F, position = "center")

# #second round - the balance table looks the same as the first round
# match_info_1 <- MatchBalance(match.out = match_model_2,
#                                     formul = balance_formula,
#                                     data = match_data_1,
#                                     print.level = 0, nboots = 1000)
# 
# match_info_1 <- MatchBalance(balance_formula,
#                                     data = match_data_2,
#                                     print.level = 0, nboots = 1000)
# 
# balance_test_after_1 <- baltest.collect(match_info_1, var.names = covariates, after = FALSE)
# balance_test_after_1 <- balance_test_after[, c("mean.Tr", "mean.Co", "T pval", "KS pval")]
# balance_test_after_1 %>%
#   kable(booktabs = T, digits = 3, align = rep('c', 4),
#         col.names = c("Treatment Mean", "Control Mean", "T p-value", "KS p-value")) %>% 
#   kable_styling(full_width = F, position = "center")
```




#### 3.1.4 MatchIt

1-on-1 nearest neighbor matching on the scaled euclidian distance, with replacement

```{r matchit version}
# https://kosukeimai.github.io/MatchIt/articles/matching-methods.html
# https://kosukeimai.github.io/MatchIt/articles/estimating-effects.html#the-standard-case


match_model_2 <- matchit(balance_formula, 
                         data = estimation_nona, 
                         estimand = "ATT",
                         method = "nearest", 
                         ratio = 1,
                         distance = "scaled_euclidean", #matching using the same distance measure as Match()
                         replace = TRUE)

summary(match_model_2) #two tables: balance before and after matching

bal.tab(match_model_2, stats = c("mean.diffs", "ks.statistics"))

#gets new dataset of just the matched observations
match_data_2 <- match_data(match_model_2) 
table(match_data_2$treat_dicho) # 849 obs: 640 treated, 209 untreated


plot(summary(match_model_2), xlim = c(0, 0.7), var.order = "unmatched") #loveplot - how different the matched is (dashed is 0.05; straight is 0.1)
title(main = "Loveplot of covariate distance")

plot(match_model_2, type = "density", interactive = FALSE) #densities of every covariate

#naive
lm_match2 <- lm_robust(trust_plebi_yes ~ treat_dicho, weights = weights, data = match_data_2) #-0.0009941906

#controls
lm_match_controls2 <- lm_robust(trust_plebi_yes ~ treat_dicho + uribe_2002_proportion + log_pop + gdp_pcap_cons2005 +  gini + etv_state_pre + etv_guerr_pre + hist_land_conflict, data = match_data_2) #-0.01619626

#use cobalt: https://kosukeimai.github.io/MatchIt/articles/assessing-balance.html#assessing-balance-with-cobalt

#each covariate plot
match_data_2 |>
  dplyr::select(all_of(covariates), treat_dicho) |> #choose covariates
  reshape2::melt(all_of(covariates), id.vars = "treat_dicho", na.rm = T) |> #put them all in one column
  ggplot(aes(x = value)) +
  geom_density(aes(col = as_factor(treat_dicho)), alpha = 0.4) +
  facet_wrap(~ variable, scales = "free") +
  labs(title = paste0("PDF of Covariates by Treatment Status after matching"),
       y = "Density", col = "Treatment") +
  labs(col = "Treatment") +
  theme_minimal() +
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank())  +
  scale_color_manual(values = c("darkorange2", "orchid3"), labels = c("Control", "Treatment")) 
```

```{r remove 2}
rm(balance_test_before, balance_test_after, match_info, match_data_1, match_data_2, match_model_1, match_model_2, exact_matches)
```


### 4.2 Propensity score matching

#### 4.2.1 Propensity score balance

```{r compare propensity scores}

# plot propensity scores of treated vs control
ggplot(tibble(ps = estimation_nona$ps, 
              treatment = as_factor(ifelse(estimation_nona$treat_dicho == 1, "Treated", "Control"))),
       aes(x = ps, col = treatment)) +
  geom_density() +
  labs(title = "Distribution of propensity scores before", y = "Density", x = "Propensity Score") +
  xlim(c(0, 1)) +
  ylim(c(0, 5)) +
  theme_minimal() +
  scale_color_manual(values = c("darkorange2", "orchid3"), labels = c("Treatment", "Control")) 

```
Very different; match by propensity to get balanced units.

#### 4.2.2 Matching propensity scores
Compare units that have a comparable probability of getting treated.


```{r match propensity scores}
#match based on propensity scores
match_ps <- Match(Y = estimation_nona$trust_plebi_yes, 
                  Tr = estimation_nona$treat_dicho, 
                  X = estimation_nona$ps, 
                  M = 1, 
                  estimand = "ATT")

#get dataset of only matched observations
match_data_ps <- estimation_nona[c(match_ps$index.treated, 
                                  match_ps$index.control),]
table(match_data_ps$treat_dicho) #723 treated, 723 control
#falsos: 479 479


```


```{r calculate att 2}
match_ps$est
match_ps$se
summary(match_ps)
# Estimate...  0.045338 
# AI SE......  0.032178 
# T-stat.....  1.409 
# p.val......  0.15885 

#calculate ATT manually
mean(match_data_ps$trust_plebi_yes[match_data_ps$treat_dicho == 1]) - mean(match_data_ps$trust_plebi_yes[match_data_ps$treat_dicho == 0]) #0.03622134

#naive matching
lm_match_ps <- lm_robust(trust_plebi_yes ~ treat_dicho, data = match_data_ps) #0.03622134

#controls matching
lm_match_ps_controls <- lm_robust(trust_plebi_yes ~ treat_dicho + uribe_2002_proportion + log_pop + gdp_pcap_cons2005 +  gini + etv_state_pre + etv_guerr_pre + hist_land_conflict, data = match_data_ps) #-0.0067888631
```

```{r balance after 2}
#get balance of dataset after matching
match_balance_ps <- MatchBalance(match.out = match_ps, 
                                 formul = balance_formula, 
                                 data = estimation_nona, 
                                 print.level = 0)

balance_table_ps <- baltest.collect(match_balance_ps, var.names = covariates, after = TRUE) #collecting numbers from matchbalance
balance_table_ps <- balance_table_ps[, c("mean.Tr", "mean.Co", "T pval", "KS pval")]

balance_table_ps |> #output table
  kable(booktabs = T, digits = 3, align = rep('c', 4),
        col.names = c("Treatment Mean", "Control Mean", "T p-value", "KS p-value")) |> 
  kable_styling(full_width = F, position = "center")

#plot of propensity scores after matching
ggplot(tibble(ps = match_data_ps$ps, 
              treatment = as_factor(ifelse(match_data_ps$treat_dicho == 1, "Treated", "Control"))),
       aes(x = ps, col = treatment)) +
  geom_density() +
  labs(title = "Distribution of propensity scores after matching", y = "Density", x = "Propensity Score") +
  xlim(c(0, 1)) +
  theme_minimal() +
  scale_color_manual(values = c("darkorange2", "orchid3"), labels = c("Control", "Treatment")) 
```
The Match() propensity scores afterwards are very close, but the data has been significantly manipulated: there are many duplicates of both control and treatment


#### 4.2.3 MatchIt: 
1-on-1 nearest neighbor matching on the propensity score, with replacement

exact = ~dem_security + hist_land_conflict,
```{r matchit version 2}

#matchit version
match_balance_ps_2 <- matchit(balance_formula, #match based on propensity scores
                    data = estimation_nona, 
                    estimand = "ATT", 
                    method = "nearest", 
                    ratio = 1,
                    distance = "glm", #glm gets propensity scores - calculated in {r cleaning data} with glm as well
                    link = "logit", 
                    replace = TRUE)

#two tables - before and after matching
summary(match_balance_ps_2)

bal.tab(match_balance_ps_2, stats = c("mean.diffs", "ks.statistics"))


#get dataset of only matched observations
match_data_ps_2 <- match_data(match_balance_ps_2) #794 obs
table(match_data_ps_2$treat_dicho) #640 treated, 154 untreated
#falsos: 154 260 


plot(summary(match_balance_ps_2), xlim = c(0, 0.7), var.order = "unmatched") #loveplot showing similarity
title(main = "Loveplot of covariate distance with propensity score matching")


plot(match_balance_ps_2, type = "density", interactive = FALSE) #density plots of each covariate

#naive
lm_match_ps2 <- lm_robust(trust_plebi_yes ~ treat_dicho, weights = weights, data = match_data_ps_2) #0.04700618

#controls
lm_match_ps_controls2 <- lm_robust(trust_plebi_yes ~ treat_dicho + uribe_2002_proportion + log_pop + gdp_pcap_cons2005 +  gini + etv_state_pre + etv_guerr_pre + hist_land_conflict, weights = weights, data = match_data_ps_2) #-0.0006025587

#comparing propensity scores after matching
ggplot(tibble(ps = match_data_ps_2$distance, 
              treatment = as_factor(ifelse(match_data_ps_2$treat_dicho == 1, "Treated", "Control"))),
       aes(x = ps, col = treatment)) +
  geom_density() +
  labs(title = "Distribution of propensity scores after matching", y = "Density", x = "Propensity Score") +
  xlim(c(0, 1)) +
  ylim(c(0, 5)) +
  theme_minimal()  +
  scale_color_manual(values = c("darkorange2", "orchid3"), labels = c("Control", "Treatment")) 
```
Some matched are worse than all for jep
Matching generally looks better with cnmh than jep


# 5. Weighting

## 5.1 Weighting propensity scores
Inverse Probability-of-Treatment Weighting (IPW)

```{r weight propensity scores}
# stabilized IPW weights, created with propensity scores
ps = estimation_nona$ps
D = estimation_nona$treat_dicho
PrD = mean(D)
IPW = (D*PrD+(1-D)*(1-PrD))/(D*ps+(1-D)*(1-ps))

# check the propensity scores before weighting
plot(density(ps[D==1]), lwd = 2, ylim = c(0, 4.5), 
     main="Distribution of propensity scores: before weighting")
lines(density(ps[D==0]),lwd = 2, lty = 2)
legend("topleft", legend = c("treated","controls"), lty = c(1, 2), lwd = 2)

# check the propensity scores after weighting
plot(density(ps[D==1], weight = IPW[D==1]/sum(IPW[D==1])), lwd = 2, ylim = c(0, 4.5), 
     main="Distribution of propensity scores: Weighted")
lines(density(ps[D==0], weight = IPW[D==0]/sum(IPW[D==0])),lwd = 2, lty = 2)
legend("topleft", legend = c("treated","controls"), lty = c(1, 2), lwd = 2)

IPW_df <- as.data.frame(IPW)

IPW_df |>
  ggplot(aes(x = IPW)) +
    geom_density() +
    labs(
    title = "Density plot of IPW",
    x = "weight"
  ) +
  theme_minimal() #check IPW for extremes 

```
todo: check how extreme the weights get / how few units are doing most of the work - it means large bias


```{r regression with weights}
#naive matching
lm_weight_ps <- lm_robust(trust_plebi_yes ~ treat_dicho, weight = IPW, data = estimation_nona) #0.04446162

#controls matching
lm_weight_ps_controls <- lm_robust(trust_plebi_yes ~ treat_dicho + uribe_2002_proportion + log_pop + gdp_pcap_cons2005 +  gini + etv_state_pre + etv_guerr_pre + hist_land_conflict, weight = IPW, data = estimation_nona) #-0.005490477
```

```{r regression with weights - WeightIt version}

#weightit version
weight_model_ps <- weightit(balance_formula, 
                     data = estimation_nona, 
                     method = "glm", 
                     estimand = "ATE")
summary(weight_model_ps)

#treatment effect
summary(lm_weightit(trust_plebi_yes ~ treat_dicho, data = estimation_nona, weightit = weight_model_ps)) #0.04446

```
TODO: These are negative and statisitically significant for most except jep - include?


## 5.2 Kernel balancing weights

```{r kernel}

#Kbal at new defaults: 24 dimensions of k; -0.01064355
weight_model_kbal = kbal(allx = estimation_nona[, covariates],
                         treatment = estimation_nona$treat_dicho,
                         printprogress = TRUE)

#naive
lm_weight_ker <- lm_robust(trust_plebi_yes ~ treat_dicho, weights = weight_model_kbal$w, data = estimation_nona)

#controls
lm_weight_ker_controls <- lm_robust(trust_plebi_yes ~ treat_dicho + uribe_2002_proportion + log_pop + gdp_pcap_cons2005 + gini + etv_state_pre + etv_guerr_pre + hist_land_conflict, weights = weight_model_kbal$w, data = estimation_nona)

# check the propensity scores before weighting
plot(density(ps[D==1]), lwd = 2, ylim = c(0, 4.5), 
     main="Distribution of propensity scores: before weighting")
lines(density(ps[D==0]),lwd = 2, lty = 2)
legend("topleft", legend = c("treated","controls"), lty = c(1, 2), lwd = 2)

# check the propensity scores after weighting
plot(density(ps[D==1], weight = weight_model_kbal$w[D==1]/sum(weight_model_kbal$w[D==1])), lwd = 2, ylim = c(0, 4.5), 
     main="Distribution of propensity scores: Kernel weighted")
lines(density(ps[D==0], weight = weight_model_kbal$w[D==0]/sum(weight_model_kbal$w[D==0])),lwd = 2, lty = 2)
legend("topleft", legend = c("treated","controls"), lty = c(1, 2), lwd = 2)

# Examine bias bound due to remaining imbalances; 
# gamma is 55
# https://cran.r-project.org/web/packages/KRLS/KRLS.pdf

#shows level of bias bound at each K; kbal() chooses where it is the lowest
plot(x = weight_model_kbal$dist_record[1,],
     y = sd(estimation_nona$trust_plebi_yes) * sqrt(55) * weight_model_kbal$dist_record[2,],
     ylab="biasbound", 
     xlab="Num. dims of K balanced",
     pch=16,
     main = "Dimensons of K and biasbound")

ker_df <- as.data.frame(weight_model_kbal[["w"]])

IPW_df |>
  ggplot(aes(x = weight_model_kbal[["w"]])) +
    geom_density() +
    labs(
    title = "Density plot of kernel weights",
    x = "weight"
  ) +
  theme_minimal()  +
  scale_color_manual(values = c("darkorange2", "orchid3"), labels = c("Control", "Treatment")) 

# 19 dimensions of K

```
above ones compare means; kernel compares means, and also x/2, and x squared, etc etc; each one of these is a dimension of K; kbal finds the dimensions of k that makes the treatment and control the most balanced (ebalance convergence); we want bias bound to be as low as possible. The plot shows the number of K and what bias bound it gives; in this case, 24 is the lowest: "Disregarding ebalance convergence and re-running at optimal choice of numdims, 24"

```{r remove 3}
rm(balance_table_ps, match_ps, match_balance_ps, match_balance_ps_2, match_data_ps, match_data_ps_2, weight_model_kbal, weight_model_ps, D, IPW, PrD, ps)
```

# 6. Results

```{r all models table - with control and naive}

#all the models
models <- list(
  "(1)" = lm_naive,
  "(2)" = lm_controls,
  "(3)" = lm_match2,
  "(4)" = lm_match_controls2,
  "(5)" = lm_match_ps2,
  "(6)" = lm_match_ps_controls2,
  "(7)" = lm_weight_ps,
  "(8)" = lm_weight_ps_controls,
  "(9)" = lm_weight_ker,
  "(10)" = lm_weight_ker_controls
  )

covariate_row <- tribble(~term, ~`(1)`, ~`(2)`, ~`(3)`, ~`(4)`, ~`(5)`, ~`(6)`, ~`(7)`, ~`(8)`, ~`(9)`, ~`(10)`,
                'controls included', 'No', 'Yes', 'No', 'Yes', 'No', 'Yes', 'No', 'Yes', 'No', 'Yes')


#put together
modelsummary(models, 
             #vcov = "HC2",
             #statistic = "p.value",
             stars = TRUE,
             #coef_omit = "uribe_2002_proportion|log_pop|gdp_pcap_cons2005|gini|etv_state_pre|etv_nonstate_pre|hist_land_conflict",
             gof_omit = "BIC|AIC|Log.Lik|F",
             add_rows = covariate_row,
             title = "The effect of Falsos Positivos cases on support for peace",
             escape = FALSE, #to fix bug in header; need to include "+ p < 0.1, * p < 0.05, ** p < 0.01, *** p < 0.001" somewhere
             output = "kableExtra") |> 
  kable_styling(full_width = TRUE, position = "center", font_size = 8)|>
  add_header_above(c(" " = 1, "Naive" = 1, "Controls" = 1, "Covariates" = 2, "Propensity scores" = 4, "Kernel" = 2)) |>
  add_header_above(c(" " = 3, "Matching" = 4, "Weighting" = 4)) #|>
  #save_kable("code/outputs/falsos_table.png")


#all five treatments:
#jep_table, jep_all_table, cnmh_table, cnmh_all_table, falsos_table
```

```{r all models plot}

models_plot <- list(
  "naive" = lm_naive,
  "controls" = lm_controls,
  " " = lm_match2,
  "covariate matching" = lm_match_controls2,
  " " = lm_match_ps2,
  "propensity score matching" = lm_match_ps_controls2,
  " " = lm_weight_ps,
  "propensity score weighting" = lm_weight_ps_controls,
  " " = lm_weight_ker,
  "kernel weighting" = lm_weight_ker_controls
  )

#plot of just treat_dicho variable
modelplot(models_plot,
          facet=TRUE,
          coef_omit = c(-2)) +
  aes(color = model, shape = model) +
  scale_color_manual(values = c("grey", "grey", "darkorange2","darkorange2","orange1", "orange1", "slateblue2", "slateblue2", "slateblue4", "slateblue4")) +
  scale_shape_manual(values = c(1, 15, 1, 15, 1, 15, 1, 15, 1, 15)) +
  theme(legend.position="none") +
  geom_vline(xintercept = 0) +
  labs(
    title = "The effect of Falsos Positivos cases on support for peace",
    x = "coefficient estimates of proportion of yes votes per municipality",
    y = "dichotomized Falsos Positivos cases",
    caption = "open circles: without controls; closed square: with controls"
  )



```

JEP:
- Adding controls makes it negative
- Propensity scores and kernel does not work well

cnmh:
- Adding controls makes it negative
- propensity score matching and weighting looks good!

Falsos:
even worse overall

# 7. Sensitivity analysis
Hopefully using uribe 2002 as benchmark? gini?

```{r sensitivity}

#The previous estimate requires the assumption of no unobserved confounders for unbiasedness.
estimation_model <- lm(trust_plebi_yes ~ treat_dicho + uribe_2002_proportion + log_pop + gdp_pcap_cons2005 + gini + etv_state_pre + etv_guerr_pre + hist_land_conflict, data = estimation_nona)


estimation_sensitivity <- sensemakr(model = estimation_model, 
                                treatment = "treat_dicho",
                                benchmark_covariates = "gini",
                                kd = c(1:3), #1-3 times stronger than treatment
                                ky = c(1:3), #1-3 times stronger than outcome
                                q = 1, #reduction of 100% of current estimate - true effect of zero is problematic
                                alpha = 0.05, 
                                reduce = T)
#estimation_sensitivity

ovb_minimal_reporting(estimation_sensitivity, format = "latex")

summary(estimation_sensitivity)

plot(estimation_sensitivity)
```
sensitivity with ols only; ipw/matching weights mess it up

```{r remove 4}
rm(estimation_model, estimation_sensitivity)
```



# 3. Falsos Positivos

## 3.1 Data cleaning 

```{r treat_definition_fp}
#select treatment
#estimation_data$treat <- estimation_data$etv_dh
#estimation_data$treat <- estimation_data$etv_all_jep
#estimation_data$treat <- estimation_data$etv_cnmh
#estimation_data$treat <- estimation_data$etv_all_cnmh
estimation_data$treat <- estimation_data$etv_false_pos

#dichotomized treatment
estimation_data <- estimation_data |>
  mutate(treat_dicho = ifelse(treat > 0, 1, 0))

#missings
#TODO: NA's found in data input.  IT IS HIGHLY RECOMMENDED THAT YOU TEST IF THE MISSING VALUES ARE BALANCED INSTEAD OF JUST DELETING THEM.

estimation_data |>
  summarise_all(~ sum(is.na(.))) #how many missing for each

estimation_nona <- estimation_data |>
  drop_na() #1115 obs to 1006 - which ones?

#covariates
covariates <- c("uribe_2002_proportion", "log_pop", "gdp_pcap_cons2005", "gini", "etv_state_pre", "etv_guerr_pre", "hist_land_conflict")
#dropped: "uribe_2006_proportion", "avg_schooling", "literacy", "prop_rural", "mpi", "dem_security",
#all: covariates <- c("uribe_2002_proportion", "uribe_2006_proportion", "avg_schooling", "literacy", "log_pop", "prop_rural", "gdp_pcap_cons2005", "mpi", "gini", "etv_state_pre", "etv_nonstate_pre", "dem_security", "hist_land_conflict")


balance_formula <- as.formula(paste("treat_dicho ~ ", paste(covariates, collapse=" + ")))

# calculate propensity scores
estimation_nona$ps <- stats::glm(balance_formula, 
                                 data = estimation_nona, 
                                 family = binomial(link = logit)
                                 )$fitted.values
```

## 2.2 Naive regression

Estimated treatment effect:
```{r naive_fp}
#naive <- lm_robust(trust_plebi_yes ~ treat, data = estimation_nona) #continuous
lm_naive <- lm_robust(trust_plebi_yes ~ treat_dicho, data = estimation_nona) #dichotomized

```


## 2.3 Naive + Control regression

```{r controls_fp}

#lm_robust(trust_plebi_yes ~ treat + uribe_2002_proportion + uribe_2006_proportion + avg_schooling + literacy + log_pop + prop_rural + gdp_pcap_cons2005 + mpi + gini + etv_state_pre + etv_nonstate_pre + dem_security + hist_land_conflict, data = estimation_nona) #continuous

lm_controls <- lm_robust(trust_plebi_yes ~ treat_dicho + uribe_2002_proportion + log_pop + gdp_pcap_cons2005 +  gini + etv_state_pre + etv_guerr_pre + hist_land_conflict, data = estimation_nona) #dicotomized
#all: trust_plebi_yes ~ treat_dicho + uribe_2002_proportion + uribe_2006_proportion + avg_schooling + literacy + log_pop + prop_rural + gdp_pcap_cons2005 + mpi + gini + etv_state_pre + etv_nonstate_pre + dem_security + hist_land_conflict, data = estimation_nona
```


## 2.4. Matching

### 3.1 Covariates

```{r treatment cutoff_fp}
estimation_nona |>
  ggplot(aes(x = treat)) + #spread of etv proportion
    geom_density() +
  labs(title = "Distribution of number of incidents per municipality")

table(estimation_nona$treat)[1] #number of zero state violence
table(estimation_nona$treat)[[1]]/nrow(estimation_nona) #percent of municipalities that have not experienced any state violence
```


#### 3.1.1 Covariate balance
```{r balance plot_fp}

estimation_nona |>
  dplyr::select(all_of(covariates), treat_dicho) |> #choose covariates
  reshape2::melt(all_of(covariates), id.vars = "treat_dicho", na.rm = T) |> #put them all in one column
  ggplot(aes(x = value)) +
  geom_density(aes(col = as_factor(treat_dicho)), alpha = 0.4) +
  facet_wrap(~ variable, scales = "free") +
  labs(title = paste0("PDF of Covariates by Treatment Status"),
       y = "Density", col = "Treatment") +
  labs(col = "Treatment") +
  theme_minimal() +
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank()) +
  scale_color_manual(values = c("darkorange2", "orchid3"), labels = c("Control", "Treatment")) 
```


```{r balance table_fp}

match_info <- MatchBalance(balance_formula, data = estimation_nona, nboots = 1000, print.level = 0) #checking balance before matching

balance_test_before <- baltest.collect(match_info, var.names = covariates, after = FALSE) #gets numbers from match_info
balance_test_before <- balance_test_before[, c("mean.Tr", "mean.Co", "T pval", "KS pval")] #rename columns

balance_test_before |> #output table
  kable(booktabs = T, digits = 3, align = rep('c', 4),
        col.names = c("Treatment Mean", "Control Mean", "T p-value", "KS p-value")) |>
  kable_styling(full_width = F, position = "center")
```

These are almost all significantly different between treatment and control.

#### 3.1.2 Matching covariates

Matching throws away non-matched control units

-Replacement = False
- how many matches: one to one


TODO: Why ATT? We are grabbing the control unit with the closest $X_i$
What does closest mean? For Match(): normalized Euclidean distance

curse of dimensionality: the more covariates, the further
  Todo: should we think through each covariate and whether we really need it?
    Also: issue of overspecifying the model
    - run model treatment ~ covariates
  
Issue of discrepancy with multiple continuous variables
  Have to adjust for it (bias correction) through BiasAdjust = TRUE
  Subtracts off estimated discrepancy
  
TODO: Different combinations: how many matches; which algorithm (Coarse - too many covariates means too far distance; pre-processing - match then regression on only close); whether to drop treatment units (removes bias)

caliper: removes units with no common support


```{r matching_fp}
exact_matches <- ifelse(colnames(estimation_nona[,covariates]) %in% c("dem_security", "hist_land_conflict"), TRUE, FALSE) #distinguish between continuous and categorical (which you can exact match on)

match_model_1 <- Match(Y = estimation_nona$trust_plebi_yes,
                       Tr = estimation_nona$treat_dicho,
                       X = estimation_nona[, covariates],
                       M = 1, # one-to-one matching
                       exact = exact_matches,
                       caliper = 2, #0.25 standard deviations dropped
                       Weight = 2, #inverse of variance
                       replace = TRUE,
                       estimand = "ATT", 
                       BiasAdjust = TRUE) # bias adjustment

#get dataset of just matched observations
match_data_1 <- estimation_nona[c(match_model_1$index.treated, 
                                  match_model_1$index.control),]
table(match_data_1$treat_dicho) #596 treated, 596 control

# #run again - doesn't make sense to do
# match_model_2 <- Match(Y = match_data_1$trust_plebi_yes,
#                        Tr = match_data_1$treat_dicho,
#                        X = match_data_1[, covariates],
#                        M = 1, # one-to-one matching
#                        estimand = "ATT", 
#                        BiasAdjust = TRUE) # bias adjustment
# 
# match_data_2 <- estimation_nona[c(match_model_2$index.treated, 
#                                   match_model_2$index.control),]
# 
# match_data_2 <- match_data_2 |>
#   drop_na()

```

```{r calculate att_fp}
match_model_1$est
match_model_1$se
summary(match_model_1)
# Estimate...  -0.011164 
# AI SE......  0.0077703 
# T-stat.....  -1.4367 
# p.val......  0.1508

#calculate ATT manually
mean(match_data_1$trust_plebi_yes[match_data_1$treat_dicho == 1]) - mean(match_data_1$trust_plebi_yes[match_data_1$treat_dicho == 0]) #0.005162898

#naive matching
#lm_match <- lm_robust(trust_plebi_yes ~ treat_dicho, data = match_data_1) #0.005162898

#controls matching
#lm_match_controls <- lm_robust(trust_plebi_yes ~ treat_dicho + uribe_2002_proportion + log_pop + gdp_pcap_cons2005 +  gini + etv_state_pre + etv_nonstate_pre + hist_land_conflict, data = match_data_1) #-0.0126571514

```


#### 3.1.3 Checking the balance after the matching

```{r balance after_fp}
match_info <- MatchBalance(match.out = match_model_1, #same as above - gets balance, but this time comparesbefore and after
                                    formul = balance_formula,
                                    data = estimation_nona,
                                    print.level = 0, nboots = 1000)

balance_test_after <- baltest.collect(match_info, var.names = covariates, after = TRUE) #new balance test
balance_test_after <- balance_test_after[, c("mean.Tr", "mean.Co", "T pval", "KS pval")]

balance_test_after %>%
  kable(booktabs = T, digits = 3, align = rep('c', 4),
        col.names = c("Treatment Mean", "Control Mean", "T p-value", "KS p-value")) %>% 
  kable_styling(full_width = F, position = "center")

# #second round - the balance table looks the same as the first round
# match_info_1 <- MatchBalance(match.out = match_model_2,
#                                     formul = balance_formula,
#                                     data = match_data_1,
#                                     print.level = 0, nboots = 1000)
# 
# match_info_1 <- MatchBalance(balance_formula,
#                                     data = match_data_2,
#                                     print.level = 0, nboots = 1000)
# 
# balance_test_after_1 <- baltest.collect(match_info_1, var.names = covariates, after = FALSE)
# balance_test_after_1 <- balance_test_after[, c("mean.Tr", "mean.Co", "T pval", "KS pval")]
# balance_test_after_1 %>%
#   kable(booktabs = T, digits = 3, align = rep('c', 4),
#         col.names = c("Treatment Mean", "Control Mean", "T p-value", "KS p-value")) %>% 
#   kable_styling(full_width = F, position = "center")
```
Balance looks better for Falsos positivos



#### 3.1.4 MatchIt

1-on-1 nearest neighbor matching on the scaled euclidian distance, without replacement

```{r matchit version_fp}
# https://kosukeimai.github.io/MatchIt/articles/matching-methods.html
# https://kosukeimai.github.io/MatchIt/articles/estimating-effects.html#the-standard-case


match_model_2 <- matchit(balance_formula, 
                         data = estimation_nona, 
                         estimand = "ATT",
                         method = "nearest", 
                         ratio = 1,
                         distance = "scaled_euclidean", #matching using the same distance measure as Match()
                         replace = TRUE)

summary(match_model_2) #two tables: balance before and after matching

bal.tab(match_model_2, stats = c("mean.diffs", "ks.statistics"))

#gets new dataset of just the matched observations
match_data_2 <- match_data(match_model_2) 
table(match_data_2$treat_dicho) # 849 obs: 640 treated, 209 untreated


plot(summary(match_model_2), xlim = c(0, 0.7), var.order = "unmatched") #loveplot - how different the matched is (dashed is 0.05; straight is 0.1)
title(main = "Loveplot of covariate distance")

plot(match_model_2, type = "density", interactive = FALSE) #densities of every covariate

#naive
lm_match2 <- lm_robust(trust_plebi_yes ~ treat_dicho, weights = weights, data = match_data_2) #-0.0009941906

#controls
lm_match_controls2 <- lm_robust(trust_plebi_yes ~ treat_dicho + uribe_2002_proportion + log_pop + gdp_pcap_cons2005 +  gini + etv_state_pre + etv_guerr_pre + hist_land_conflict, data = match_data_2) #-0.01619626

#use cobalt: https://kosukeimai.github.io/MatchIt/articles/assessing-balance.html#assessing-balance-with-cobalt

#each covariate plot
match_data_2 |>
  dplyr::select(all_of(covariates), treat_dicho) |> #choose covariates
  reshape2::melt(all_of(covariates), id.vars = "treat_dicho", na.rm = T) |> #put them all in one column
  ggplot(aes(x = value)) +
  geom_density(aes(col = as_factor(treat_dicho)), alpha = 0.4) +
  facet_wrap(~ variable, scales = "free") +
  labs(title = paste0("PDF of Covariates by Treatment Status after matching"),
       y = "Density", col = "Treatment") +
  labs(col = "Treatment") +
  theme_minimal() +
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank())  +
  scale_color_manual(values = c("darkorange2", "orchid3"), labels = c("Control", "Treatment")) 
```

```{r remove 2_fp}
rm(balance_test_before, balance_test_after, match_info, match_data_1, match_data_2, match_model_1, match_model_2, exact_matches)
```


### 4.2 Propensity score matching

#### 4.2.1 Propensity score balance

```{r compare propensity scores_fp}

# plot propensity scores of treated vs control
ggplot(tibble(ps = estimation_nona$ps, 
              treatment = as_factor(ifelse(estimation_nona$treat_dicho == 1, "Treated", "Control"))),
       aes(x = ps, col = treatment)) +
  geom_density() +
  labs(title = "Distribution of propensity scores", y = "Density", x = "Propensity Score") +
  xlim(c(0, 1)) +
  ylim(c(0, 5)) +
  theme_minimal() +
  scale_color_manual(values = c("darkorange2", "orchid3"), labels = c("Treatment", "Control")) 

```
Very different; match by propensity to get balanced units.

#### 4.2.2 Matching propensity scores
Compare units that have a comparable probability of getting treated.


```{r match propensity scores_fp}
#match based on propensity scores
match_ps <- Match(Y = estimation_nona$trust_plebi_yes, 
                  Tr = estimation_nona$treat_dicho, 
                  X = estimation_nona$ps, 
                  M = 1, 
                  estimand = "ATT")

#get dataset of only matched observations
match_data_ps <- estimation_nona[c(match_ps$index.treated, 
                                  match_ps$index.control),]
table(match_data_ps$treat_dicho) #723 treated, 723 control
#falsos: 479 479


```


```{r calculate att 2_fp}
match_ps$est
match_ps$se
summary(match_ps)
# Estimate...  0.045338 
# AI SE......  0.032178 
# T-stat.....  1.409 
# p.val......  0.15885 

#calculate ATT manually
mean(match_data_ps$trust_plebi_yes[match_data_ps$treat_dicho == 1]) - mean(match_data_ps$trust_plebi_yes[match_data_ps$treat_dicho == 0]) #0.03622134

#naive matching
lm_match_ps <- lm_robust(trust_plebi_yes ~ treat_dicho, data = match_data_ps) #0.03622134

#controls matching
lm_match_ps_controls <- lm_robust(trust_plebi_yes ~ treat_dicho + uribe_2002_proportion + log_pop + gdp_pcap_cons2005 +  gini + etv_state_pre + etv_guerr_pre + hist_land_conflict, data = match_data_ps) #-0.0067888631
```

```{r balance after 2_fp}
#get balance of dataset after matching
match_balance_ps <- MatchBalance(match.out = match_ps, 
                                 formul = balance_formula, 
                                 data = estimation_nona, 
                                 print.level = 0)

balance_table_ps <- baltest.collect(match_balance_ps, var.names = covariates, after = TRUE) #collecting numbers from matchbalance
balance_table_ps <- balance_table_ps[, c("mean.Tr", "mean.Co", "T pval", "KS pval")]

balance_table_ps |> #output table
  kable(booktabs = T, digits = 3, align = rep('c', 4),
        col.names = c("Treatment Mean", "Control Mean", "T p-value", "KS p-value")) |> 
  kable_styling(full_width = F, position = "center")

#plot of propensity scores after matching
ggplot(tibble(ps = match_data_ps$ps, 
              treatment = as_factor(ifelse(match_data_ps$treat_dicho == 1, "Treated", "Control"))),
       aes(x = ps, col = treatment)) +
  geom_density() +
  labs(title = "Distribution of propensity scores after matching", y = "Density", x = "Propensity Score") +
  xlim(c(0, 1)) +
  theme_minimal() +
  scale_color_manual(values = c("darkorange2", "orchid3"), labels = c("Control", "Treatment")) 
```
The Match() propensity scores afterwards are very close, but the data has been significantly manipulated: there are many duplicates of both control and treatment


#### 4.2.3 MatchIt: 
1-on-1 nearest neighbor matching on the propensity score, with replacement

exact = ~dem_security + hist_land_conflict,
```{r matchit version 2_fp}

#matchit version
match_balance_ps_2 <- matchit(balance_formula, #match based on propensity scores
                    data = estimation_nona, 
                    estimand = "ATT", 
                    method = "nearest", 
                    ratio = 1,
                    distance = "glm", #glm gets propensity scores - calculated in {r cleaning data} with glm as well
                    link = "logit", 
                    replace = TRUE)

#two tables - before and after matching
summary(match_balance_ps_2)

bal.tab(match_balance_ps_2, stats = c("mean.diffs", "ks.statistics"))


#get dataset of only matched observations
match_data_ps_2 <- match_data(match_balance_ps_2) #794 obs
table(match_data_ps_2$treat_dicho) #640 treated, 154 untreated
#falsos: 154 260 


plot(summary(match_balance_ps_2), xlim = c(0, 0.7), var.order = "unmatched") #loveplot showing similarity
title(main = "Loveplot of covariate distance with propensity score matching")


plot(match_balance_ps_2, type = "density", interactive = FALSE) #density plots of each covariate

#naive
lm_match_ps2 <- lm_robust(trust_plebi_yes ~ treat_dicho, weights = weights, data = match_data_ps_2) #0.04700618

#controls
lm_match_ps_controls2 <- lm_robust(trust_plebi_yes ~ treat_dicho + uribe_2002_proportion + log_pop + gdp_pcap_cons2005 +  gini + etv_state_pre + etv_guerr_pre + hist_land_conflict, weights = weights, data = match_data_ps_2) #-0.0006025587

#comparing propensity scores after matching
ggplot(tibble(ps = match_data_ps_2$distance, 
              treatment = as_factor(ifelse(match_data_ps_2$treat_dicho == 1, "Treated", "Control"))),
       aes(x = ps, col = treatment)) +
  geom_density() +
  labs(title = "Distribution of propensity scores after matching", y = "Density", x = "Propensity Score") +
  xlim(c(0, 1)) +
  ylim(c(0, 5)) +
  theme_minimal()  +
  scale_color_manual(values = c("darkorange2", "orchid3"), labels = c("Control", "Treatment")) 
```
Some matched are worse than all for jep
Matching generally looks better with cnmh than jep


# 5. Weighting

## 5.1 Weighting propensity scores
Inverse Probability-of-Treatment Weighting (IPW)

```{r weight propensity scores_fp}
# stabilized IPW weights, created with propensity scores
ps = estimation_nona$ps
D = estimation_nona$treat_dicho
PrD = mean(D)
IPW = (D*PrD+(1-D)*(1-PrD))/(D*ps+(1-D)*(1-ps))

# check the propensity scores before weighting
plot(density(ps[D==1]), lwd = 2, ylim = c(0, 4.5), 
     main="Distribution of propensity scores: before weighting")
lines(density(ps[D==0]),lwd = 2, lty = 2)
legend("topleft", legend = c("treated","controls"), lty = c(1, 2), lwd = 2)

# check the propensity scores after weighting
plot(density(ps[D==1], weight = IPW[D==1]/sum(IPW[D==1])), lwd = 2, ylim = c(0, 4.5), 
     main="Distribution of propensity scores: Weighted")
lines(density(ps[D==0], weight = IPW[D==0]/sum(IPW[D==0])),lwd = 2, lty = 2)
legend("topleft", legend = c("treated","controls"), lty = c(1, 2), lwd = 2)

IPW_df <- as.data.frame(IPW)

IPW_df |>
  ggplot(aes(x = IPW)) +
    geom_density() +
    labs(
    title = "Density plot of IPW",
    x = "weight"
  ) +
  theme_minimal() #check IPW for extremes 

```
todo: check how extreme the weights get / how few units are doing most of the work - it means large bias


```{r regression with weights_fp}
#naive matching
lm_weight_ps <- lm_robust(trust_plebi_yes ~ treat_dicho, weight = IPW, data = estimation_nona) #0.04446162

#controls matching
lm_weight_ps_controls <- lm_robust(trust_plebi_yes ~ treat_dicho + uribe_2002_proportion + log_pop + gdp_pcap_cons2005 +  gini + etv_state_pre + etv_guerr_pre + hist_land_conflict, weight = IPW, data = estimation_nona) #-0.005490477
```

```{r regression with weights - WeightIt version_fp}

#weightit version
weight_model_ps <- weightit(balance_formula, 
                     data = estimation_nona, 
                     method = "glm", 
                     estimand = "ATE")
summary(weight_model_ps)

#treatment effect
summary(lm_weightit(trust_plebi_yes ~ treat_dicho, data = estimation_nona, weightit = weight_model_ps)) #0.04446

```
TODO: These are negative and statisitically significant for most except jep - include?


## 5.2 Kernel balancing weights

```{r kernel_fp}

#Kbal at new defaults: 24 dimensions of k; -0.01064355
weight_model_kbal = kbal(allx = estimation_nona[, covariates],
                         treatment = estimation_nona$treat_dicho,
                         printprogress = TRUE)

estimation_nona$ker_weights <- weight_model_kbal$w

#naive
lm_weight_ker <- lm_robust(trust_plebi_yes ~ treat_dicho, weights = weight_model_kbal$w, data = estimation_nona)

#controls
lm_weight_ker_controls <- lm_robust(trust_plebi_yes ~ treat_dicho + uribe_2002_proportion + log_pop + gdp_pcap_cons2005 + gini + etv_state_pre + etv_guerr_pre + hist_land_conflict, weights = weight_model_kbal$w, data = estimation_nona)

# check covariate balance before weighting
estimation_nona |>
  dplyr::select(all_of(covariates), treat_dicho) |> #choose covariates
  reshape2::melt(all_of(covariates), id.vars = "treat_dicho", na.rm = T) |> #put them all in one column
  ggplot(aes(x = value)) +
  geom_density(aes(col = as_factor(treat_dicho)), alpha = 0.4) +
  facet_wrap(~ variable, scales = "free") +
  labs(title = paste0("PDF of Covariates by Treatment Status"),
       y = "Density", col = "Treatment") +
  labs(col = "Treatment") +
  theme_minimal() +
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank()) +
  scale_color_manual(values = c("darkorange2", "orchid3"), labels = c("Control", "Treatment")) 

# check covariate balance after weighting
estimation_nona |>
  dplyr::select(all_of(covariates), treat_dicho, ker_weights) |> #choose covariates
  reshape2::melt(all_of(covariates), id.vars = c("treat_dicho", "ker_weights"), na.rm = T) |> #put them all in one column
  ggplot(aes(x = value)) +
  geom_density(aes(col = as_factor(treat_dicho), weight = ker_weights), alpha = 0.4) +
  facet_wrap(~ variable, scales = "free") +
  labs(title = paste0("PDF of Covariates by Treatment Status after kernel weights"),
       y = "Density", col = "Treatment") +
  labs(col = "Treatment") +
  theme_minimal() +
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank())  +
  scale_color_manual(values = c("darkorange2", "orchid3"), labels = c("Control", "Treatment")) 

aes(weight = poptotal)

# check the propensity scores before weighting
plot(density(ps[D==1]), lwd = 2, ylim = c(0, 4.5), 
     main="Distribution of propensity scores: before weighting")
lines(density(ps[D==0]),lwd = 2, lty = 2)
legend("topleft", legend = c("treated","controls"), lty = c(1, 2), lwd = 2)

# check the propensity scores after weighting
plot(density(ps[D==1], weight = weight_model_kbal$w[D==1]/sum(weight_model_kbal$w[D==1])), lwd = 2, ylim = c(0, 4.5), 
     main="Distribution of propensity scores: Kernel weighted")
lines(density(ps[D==0], weight = weight_model_kbal$w[D==0]/sum(weight_model_kbal$w[D==0])),lwd = 2, lty = 2)
legend("topleft", legend = c("treated","controls"), lty = c(1, 2), lwd = 2)

# Examine bias bound due to remaining imbalances; 
# gamma is 55
# https://cran.r-project.org/web/packages/KRLS/KRLS.pdf

#shows level of bias bound at each K; kbal() chooses where it is the lowest
plot(x = weight_model_kbal$dist_record[1,],
     y = sd(estimation_nona$trust_plebi_yes) * sqrt(55) * weight_model_kbal$dist_record[2,],
     ylab="biasbound", 
     xlab="Num. dims of K balanced",
     pch=16,
     main = "Dimensons of K and biasbound")

ker_df <- as.data.frame(weight_model_kbal[["w"]])

IPW_df |>
  ggplot(aes(x = weight_model_kbal[["w"]])) +
    geom_density() +
    labs(
    title = "Density plot of kernel weights",
    x = "weight"
  ) +
  theme_minimal()  +
  scale_color_manual(values = c("darkorange2", "orchid3"), labels = c("Control", "Treatment")) 

# 29 dimensions of K

```
above ones compare means; kernel compares means, and also x/2, and x squared, etc etc; each one of these is a dimension of K; kbal finds the dimensions of k that makes the treatment and control the most balanced (ebalance convergence); we want bias bound to be as low as possible. The plot shows the number of K and what bias bound it gives; in this case, 24 is the lowest: "Disregarding ebalance convergence and re-running at optimal choice of numdims, 24"

```{r remove 3_fp}
rm(balance_table_ps, match_ps, match_balance_ps, match_balance_ps_2, match_data_ps, match_data_ps_2, weight_model_kbal, weight_model_ps, D, IPW, PrD, ps)
```

# 6. Results

```{r just kernel}

#put together
modelsummary(lm_weight_ker_controls, 
             #vcov = "HC2",
             #statistic = "p.value",
             stars = TRUE,
             #coef_omit = "uribe_2002_proportion|log_pop|gdp_pcap_cons2005|gini|etv_state_pre|etv_nonstate_pre|hist_land_conflict",
             gof_omit = "BIC|AIC|Log.Lik|F",
             title = "The effect of Falsos Positivos cases on support for peace, kernel weighted",
             escape = FALSE, #to fix bug in header; need to include "+ p < 0.1, * p < 0.05, ** p < 0.01, *** p < 0.001" somewhere
             output = "kableExtra")
```

```{r all models table_fp - only controls}

#all the models
models <- list(
  "(1)" = lm_naive,
  "(2)" = lm_controls,
  "(3)" = lm_match_controls2,
  "(4)" = lm_match_ps_controls2,
  "(5)" = lm_weight_ps_controls,
  "(6)" = lm_weight_ker_controls
  )

covariate_row <- tribble(~term, ~`(1)`, ~`(2)`, ~`(3)`, ~`(4)`, ~`(5)`, ~`(6)`,
                'controls included', 'No', 'Yes', 'Yes', 'Yes','Yes', 'Yes')


#put together
modelsummary(models, 
             #vcov = "HC2",
             #statistic = "p.value",
             stars = TRUE,
             #coef_omit = "uribe_2002_proportion|log_pop|gdp_pcap_cons2005|gini|etv_state_pre|etv_nonstate_pre|hist_land_conflict",
             gof_omit = "BIC|AIC|Log.Lik|F",
             add_rows = covariate_row,
             title = "The effect of Falsos Positivos cases on support for peace",
             escape = FALSE, #to fix bug in header; need to include "+ p < 0.1, * p < 0.05, ** p < 0.01, *** p < 0.001" somewhere
             output = "kableExtra") |> 
  kable_styling(full_width = TRUE, position = "center", font_size = 8) |>
  add_header_above(c(" " = 1, "Naive" = 1, "Controls" = 1, "Covariates" = 1, "Propensity scores" = 2, "Kernel" = 1)) |>
  add_header_above(c(" " = 3, "Matching" = 2, "Weighting" = 2)) #|>
  #save_kable("code/outputs/falsos_table.png")


#all five treatments:
#jep_table, jep_all_table, cnmh_table, cnmh_all_table, falsos_table
```



```{r all models table_fp}

#all the models
models <- list(
  "(1)" = lm_naive,
  "(2)" = lm_controls,
  "(3)" = lm_match2,
  "(4)" = lm_match_controls2,
  "(5)" = lm_match_ps2,
  "(6)" = lm_match_ps_controls2,
  "(7)" = lm_weight_ps,
  "(8)" = lm_weight_ps_controls,
  "(9)" = lm_weight_ker,
  "(10)" = lm_weight_ker_controls
  )

covariate_row <- tribble(~term, ~`(1)`, ~`(2)`, ~`(3)`, ~`(4)`, ~`(5)`, ~`(6)`, ~`(7)`, ~`(8)`, ~`(9)`, ~`(10)`,
                'controls included', 'No', 'Yes', 'No', 'Yes', 'No', 'Yes', 'No', 'Yes', 'No', 'Yes')


#put together
modelsummary(models, 
             #vcov = "HC2",
             #statistic = "p.value",
             stars = TRUE,
             #coef_omit = "uribe_2002_proportion|log_pop|gdp_pcap_cons2005|gini|etv_state_pre|etv_nonstate_pre|hist_land_conflict",
             gof_omit = "BIC|AIC|Log.Lik|F",
             add_rows = covariate_row,
             title = "The effect of Falsos Positivos cases on support for peace",
             escape = FALSE, #to fix bug in header; need to include "+ p < 0.1, * p < 0.05, ** p < 0.01, *** p < 0.001" somewhere
             output = "kableExtra") |> 
  kable_styling(full_width = TRUE, position = "center", font_size = 8)|>
  add_header_above(c(" " = 1, "Naive" = 1, "Controls" = 1, "Covariates" = 2, "Propensity scores" = 4, "Kernel" = 2)) |>
  add_header_above(c(" " = 3, "Matching" = 4, "Weighting" = 4)) #|>
  #save_kable("code/outputs/falsos_table.png")


#all five treatments:
#jep_table, jep_all_table, cnmh_table, cnmh_all_table, falsos_table
```


```{r all models plot_fp - only controls}

models_plot <- list(
  "naive" = lm_naive,
  "controls" = lm_controls,
  "covariate matching" = lm_match_controls2,
  "propensity score matching" = lm_match_ps_controls2,
  "propensity score weighting" = lm_weight_ps_controls,
  "kernel weighting" = lm_weight_ker_controls
  )

#plot of just treat_dicho variable
modelplot(models_plot,
          facet=TRUE,
          coef_omit = c(-2)) +
  aes(color = model, shape = model) +
  scale_color_manual(values = c("grey", "grey", "darkorange2","orange1", "slateblue2", "slateblue4")) +
  scale_shape_manual(values = c(1, 15, 15, 15, 15, 15)) +
  theme(legend.position="none") +
  geom_vline(xintercept = 0) +
  labs(
    title = "The effect of Falsos Positivos cases on support for peace",
    x = "coefficient estimates of proportion of yes votes per municipality",
    y = "dichotomized Falsos Positivos cases",
    caption = "open circles: without controls; closed square: with controls"
  )



```


```{r all models plot_fp}

models_plot <- list(
  "naive" = lm_naive,
  "controls" = lm_controls,
  " " = lm_match2,
  "covariate matching" = lm_match_controls2,
  " " = lm_match_ps2,
  "propensity score matching" = lm_match_ps_controls2,
  " " = lm_weight_ps,
  "propensity score weighting" = lm_weight_ps_controls,
  " " = lm_weight_ker,
  "kernel weighting" = lm_weight_ker_controls
  )

#plot of just treat_dicho variable
modelplot(models_plot,
          facet=TRUE,
          coef_omit = c(-2)) +
  aes(color = model, shape = model) +
  scale_color_manual(values = c("grey", "grey", "darkorange2","darkorange2","orange1", "orange1", "slateblue2", "slateblue2", "slateblue4", "slateblue4")) +
  scale_shape_manual(values = c(1, 15, 1, 15, 1, 15, 1, 15, 1, 15)) +
  theme(legend.position="none") +
  geom_vline(xintercept = 0) +
  labs(
    title = "The effect of Falsos Positivos cases on support for peace",
    x = "coefficient estimates of proportion of yes votes per municipality",
    y = "dichotomized Falsos Positivos cases",
    caption = "open circles: without controls; closed square: with controls"
  )



```

JEP:
- Adding controls makes it negative
- Propensity scores and kernel does not work well

cnmh:
- Adding controls makes it negative
- propensity score matching and weighting looks good!

Falsos:
even worse overall

# 7. Sensitivity analysis
Hopefully using uribe 2002 as benchmark? gini?

```{r sensitivity_fp}

#The previous estimate requires the assumption of no unobserved confounders for unbiasedness.
estimation_model <- lm(trust_plebi_yes ~ treat_dicho + uribe_2002_proportion + log_pop + gdp_pcap_cons2005 + gini + etv_state_pre + etv_guerr_pre + hist_land_conflict, data = estimation_nona)


estimation_sensitivity <- sensemakr(model = estimation_model, 
                                treatment = "treat_dicho",
                                benchmark_covariates = "gini",
                                kd = c(1:3), #1-3 times stronger than treatment
                                ky = c(1:3), #1-3 times stronger than outcome
                                q = 1, #reduction of 100% of current estimate - true effect of zero is problematic
                                alpha = 0.05, 
                                reduce = T)
#estimation_sensitivity

ovb_minimal_reporting(estimation_sensitivity, format = "latex")

summary(estimation_sensitivity)

plot(estimation_sensitivity,
     main = "Sensitivity analysis with gini as benchmark")
```
sensitivity with ols only; ipw/matching weights mess it up

```{r remove 4_fp}
rm(estimation_model, estimation_sensitivity)
```

  

# 8. Robustness checks:
  Falsos Positivos, State + Paramilitaries, And any kind of violence
  Changing time





